### 2009-04-18 response-to-reviews.txt
Import from CVS.

> ------------------------------------------------------------------------
> 
> Reviewer #1 (Remarks for the Author):

Both reviews were extremely constructive and a genuine pleasure to
read.

> There are very interesting and potentially quite influential ideas in
> this paper, but too many of them at once are breathlessly muddled up
> together in a fog of words, without sufficiently careful description
> and logical analysis.

We are excited about the possibilities of this project, which we began
3 years ago with the goal of building algorithms to reconstruct
ancient ribosomal RNA.

The main products of this work, and the points of the paper, are:

 1. the mathematics of the new algorithms (tree transducer composition
 and "loopy DP");

 2. implementations of these algorithms as working code (the
 "indiegram" program).

The investigations of various other aspects of RNA reconstruction
(such as error rates, alignment quality & benchmarks of our other
software programs) are secondary to these main algorithmic points.

We agree that the inclusion of these secondary aspects in the main
paper was distracting. Accordingly we have moved these sections to the
supplementary information and trimmed them substantially.

(See below for an explanation of why they were included in the first
place, and what their relevance is; essentially, they are precursors to
future benchmarks we anticipate conducting pending implementation of
more sophisticated MCMC kernel algorithms.)

> The paper does not convey a clear, thorough, reproducible piece of
> science to the reader. It lacks focus, discipline, and rigor. The
> authors need to choose one story they want to tell, then tell it
> completely and thoroughly.

The story we are trying to tell is the systematic extension of
Pair-SCFG-like models to multiple sequences, with a working (albeit
memory-hungry) implementation on a three-branch tree.

We have attempted to bring this story into much sharper focus with our
revised Introduction and structure.

> The paper opens with some grand statements about a research program to
> develop "paleogenetic" algorithms to reconstruct the RNA world. I
> think the authors are using this story to dress up what they're doing,
> rather than really taking this research program seriously and
> analyzing it logically. For starters, in order to phylogenetically
> reconstruct a "primordial ribozyme" such as the "all-RNA primordial
> ribosome" from the "RNA world" from extant sequences, the authors need
> to be able to argue that the last common ancestor of rRNAs lived in
> the RNA world, not the modern DNA/RNA/protein world. It seems clear,
> however, that the last universal common ancestor (LUCA) of all known
> extant life had a reasonably complicated and more or less standard
> DNA/RNA/protein biochemistry, because (even taking abundant lateral
> transfer into account) of how many proteins and biochemical pathways
> can be traced back to LUCA. A serious attempt to reconstruct any
> aspect of the RNA world by phylogenetic analysis must deal with the
> probability that LUCA is almost certainly a bottleneck that occurs
> after the RNA world had passed (if the RNA world existed at all). This
> is not to say that ancestral reconstruction of structural RNAs isn't
> interesting, even if they reconstruct RNAs that lived in less fabulous
> creatures than ribosaurs -- it's only to say that the authors' stated
> goals seem under-analyzed and over-grandiose, and therefore are
> offputting to the reader. Much of the introduction's verbiage about
> the RNA world could be deleted, and the aims of the paper could be
> focused more realistically, concisely, and clearly.

Point taken.

Reconstructing early ribosomes has been a goal of this work for
several years, and Crick's all-RNA ribosome is often cited as a
convincing argument (or at least an inspiration) for the RNA world
hypothesis. So we did not think that, by quoting the RNA world
literature, we were being grandiose or trying to "dress up" our work.

However, we see the reviewer's point that "LUCA != RNA world" and we
agree it would be clearer and more consistent to focus on LUCA.  (Our
algorithms can only reconstruct RNA sequence from periods during which
a "more or less standard biochemistry" was established, and the
reviewers' points about bottlenecks at LUCA are good ones.)

Accordingly, we have rewritten the introduction with more realistic
expectations for the epochs we may actually be able to reach (in fact,
only the first paragraphs of the Introduction and Abstract ever
mentioned the RNA World; the newly revised manuscript barely mentions
it).

We enjoyed the word "ribosaur" -- is it published, or otherwise in
current usage? (Not according to Google...)

We think that the word "paleogenetics" is reasonably well-established
(by the work of, e.g. Pauling & Zuckerkandl, Svante Paabo, Steve
Benner, etc., which we think this naturally relates to); however, if
the reviewer feels that it is confusing or inappropriate, then we can
change the wording accordingly.

> At the same time, the introduction lacks a clear problem definition
> statement. What is the input; what is the desired output; what is the
> main idea behind the approach for transforming input to output?

The central problem, as we now state more clearly in the Introduction,
is this:

 Given a conditionally normalized Pair SCFG, modeling P(Y|X) where X
 and Y are foldback RNA structures, along with a phylogenetic tree,
 what is the corresponding multiple-sequence SCFG; and can we run the
 Cocke-Younger-Kasami algorithm on it (suitably constrained) ?

The solution to this problem is given by the transducer composition
algorithm, which is the core result of this paper.

> Throughout the paper, terms and programs are thrown around in a
> muddled soup, as if everything is a general purpose magic inference
> tool. _xrate_ estimates substitution rate parameters (p.2) and folds
> RNA (p.5); _evoldoer_ generates pair SCFGs (p.3) and "implements"
> TKFST (p.5), whatever "implements" may mean (any given probabilistic
> model may be used for a wide range of different simulations or
> inference problems). Do they slice and dice tomatoes too? At far too
> many points, the authors recite lists of possible applications that
> might be tackled someday, or lists of other approaches that can be
> viewed as special cases of their approach -- without ever particularly
> clearly defining what their approach is. I'm all in favor of
> generalizable approaches and powerful probabilistic models, but I want
> to see at least one specific, well-described problem solved before one
> expounds on all the other things that might possibly be done. That's
> what the Discussion is for.

We have emphasized the central problem of constructing a model of
multiple sequences given a model of the evolution of two sequences and
a phylogenetic tree (the "transducer composition" algorithm) in the
revised manuscript, including a full statement of inputs and outputs.

We have also removed excessive references to program names as
requested by both reviewers.

We have attempted to clarify the possible applications of these
methodologies and place them in the context of prior work.  Some of
these applications/questions are:

-- Given a conditional distribution P(Y | X), can one invert this and
compute (or sample from) the distribution of the ancestral sequence,
i.e. P(X | Y)?

-- Can one repeatedly apply this model and e.g. sample Z from P(Z|Y)
where Z is also a sequence/structure? Can one sum over all values of Y
and compute P(Z|X) directly, or sample Z from it?  (This is indirectly
related to the problem of time-dependence, since the branch from X-->Z
represents twice as much evolutionary time as the branch from X-->Y.)

-- How well does the model handle multiple sequences and missing data?
E.g. can we sample from the distribution of unobserved "grandparent"
sequences, given a set of observed "grandchildren" (and unobserved
"parents")?

-- What about if you have an unobserved parent with three observed
children? Can you sample from P(parent|children) ?  (This is closest
to the task we actually implement; doing this is sufficient to address
all of the other questions using an MCMC approach, as shown by our
2001 paper and several other researchers.)

-- Can we estimate the multiple alignment and phylogeny of a given
set of sequences, starting only with the model P(Y | X) ?

All of these questions can be addressed by the multi-sequence SCFGs we
describe, using standard SCFG algorithms (Inside-Outside, CYK, etc.)
and constrained MCMC variants on those algorithms.

> The authors could do a much better job of recognizing that ancestral
> sequence reconstruction has been done for decades: both parsimony
> methods and maximum likelihood methods for phylogenetic inference
> explicitly infer sequences at ancestral nodes in a tree.

Thank you for pointing this out; we did not intend to slight previous
authors.  In our rewritten Discussion, we have cited more of the
basic papers that established the computational and mathematical
foundations of the field.  We also cite the 2007 book edited by David
Liberles, an excellent survey of current uses of substitution models
for ancestral reconstruction, as well as the origins of the field
(e.g. Pauling & Zuckerkandl's paper, parsimony & ML methods, and so on).

> What's of interest here is whether that can be done not only for
> residue substitution events, but for insertion/deletion events as
> well,

We completely agree.
Insertions and deletions in unstructured sequences can be well-handled by
conditionally-normalized Pair HMMs, or "string transducer" state machines,
for which several implementations of the transducer composition
algorithm for building multi-sequence models from pairwise models now exist.
One of the central goals of this paper was to extend this theory to
the case of RNA secondary structures (or parse-tree transducers), in
order to model effects like indels that conserve RNA structure.

> and whether the inferences are sufficiently accurate that we can
> take them seriously and actually synthesize the inferred ancestor(s)
> and believe that they tell us something about the structure and
> function of extinct biomolecules.
> 
> The Methods are unnecessarily opaque and confusing, focusing too much
> on peripheral terminology and too little on an actual problem at
> hand. For example, for all this talk of "transducer theory", it is
> never clear how a "transducer" is meaningfully different from any
> other conditional probability model used in sequence analysis -- such
> as profile HMMs and profile SCFGs, which model a conditional
> distribution P(Y | X) via a model built from sequence X, P(Y | X) ~=
> P(Y | model) P(model | X) using a point estimate of an appropriate
> model derived from X, where X could be either a query
> sequence/alignment in a similarity search,

We apologize for the lack of clarity.  We have attempted to rectify
this in the rewritten manuscript by first discussing
jointly-normalized models (Pair HMMs) and then introducing
conditionally-normalized models (transducers).

We believe that the transducer is a meaningful improvement on
HMMs/SCFGs because of the transducer composition algorithm, which
offers a natural way to construct models of multiple sequences; for
example, given a model P(Y|X), and a suitable phylogeny, we can build
a model like this

   P(A, B, C, D, E | X, TREE)

where X is the root sequence of the tree, and {A,B,C,D,E} are leaves.
This model correctly accounts for all combinations of overlapping
indel events that can possibly occur on the branches of the tree --
for example, an insertion on an ancient branch can be interrupted by
an insertion on an intermediate branch, then both insertions can be
[partially] deleted on an even younger branch. The model is a
multi-sequence HMM, so that multiple sequence alignment is simply an
application of the Viterbi algorithm (or posterior decoding), and MCMC
Gibbs samplers can be constructed by applying suitable constraints to
the Forward algorithm.

Furthermore, dynamic programming (and MCMC-sampling) algorithms exist
to invert this model

   P(X | A, B, C, D, E, TREE)

Ancestral reconstruction involves sampling from and/or maximizing this
inverted distribution. Many other Bayesian manipulations can similarly
be performed on this likelihood function.

This ability to generalize systematically to multiple sequences is
missing from HMMs and SCFGs. While it is possible to convert a
single-sequence HMM into a phylo-HMM (by having it emit alignment
columns instead of single symbols), and even to simulate gaps to some
extent (by modeling them as an additional character), there is no
simple recipe (other than transducer composition) for extending a Pair
HMM to a corresponding multi-sequence HMM that includes affine gap
penalties, overlapping indels, nested indels, etc.

> model derived from X, where X could be [...]
> an ancestor on a tree in a phylogenetic inference problem.

We appreciate the point that there are similarities between Profile
SCFGs and the transducers which we describe here. For example, the
paper by Klein & Eddy (2003), describing the RSEARCH program,
introduces a profile SCFG which may be regarded as modeling a
conditional distribution P(Y|X). However, to our knowledge, existing
literature on Profile SCFGs does not describe the formalisms which we
need to reconstruct structured sequences on trees.  In particular, we
need ways to find:

-- the corresponding joint Pair-SCFG, modeling P(X, Y);
-- the Bayesian inverse Profile-SCFG, modeling P(X | Y);
-- situations involving more than two sequences, e.g. the
   Profile-SCFG for P(X | Y, Z), where X is the last common ancestor
   of Y and Z; or the corresponding Pair-SCFG for P(Y, Z); or the
   corresponding Triplet-SCFG for P(X, Y, Z).

Transducers provide a generic, systematic way to build such
multi-sequence, phylogenetically-structured SCFGs. In this paper we
attempt to describe how to perform these procedures.

> Seems to me that the critical thing,
> which the authors actually _underemphasize_, is how one makes this
> conditional probability model P(Y | X) be dependent on evolutionary
> time t, P(Y | X,t), via an underlying continuous-time Markov model of
> evolution in terms of rates of individual events including
> substitutions, insertions, and deletions.

This is a good point.  Time-dependent parameterization is indeed
critical, but it can be handled as a separate problem from the one we
address here.  Our goal is to provide a modular solution to the
ancestral reconstruction problem, wherein the problem of creating
accurate (two-sequence) models can be treated as disjoint from the
problem of extending those models to multiple sequences.  In this
paper we have focused on the second problem: Given a model of the
evolution of two sequences, how do we construct a model of many
sequences?

For this reason, while the time parametrization problem is certainly
central to the ancestral reconstruction problem, we believe that it
can be treated as a distinct problem from the multi-sequence model
construction problem which is the focus of this paper.  As the
reviewer points out, we largely gloss over how to choose an
appropriate time parametrization, and instead focus on:
Suppose you do have such a time-dependent model. How do you then
extend it from two sequences to multiple sequences?

More generally, how do you sample from some generic distribution

P(X_1, X_2 ... X_M  |  Y_1, Y_2 ... Y_M,   TREE)

where {X_1 ... X_M} are all ancestors, and {Y_1 ... Y_M} are all
descendants, related by a given TREE?

For substitution models, the answer to this question involves
Felsenstein's pruning algorithm.

For indel/RNA models, our answer is that you construct a
multi-sequence SCFG and then sample paths through it (appropriately
conditioned on the data {Y_n}).  Transducers allow you to build this
Multi-SCFG, assuming you already know the Pair-SCFG (i.e. the
parse-tree transducer).

Note that neither of these answers is the same as the problem of
getting time-dependent parameters! For substitution processes, you get
time-dependent parameters by matrix exponentiation (i.e. solving a
continuous-time Markov chain).

For RNA/indel processes, the TKFST model described in earlier work
(Holmes 2004) represents one attempt to get to these time-dependent
parameters for point substitutions, covariant substitutions,
insertions, deletions and structural changes.

TKFST is flawed, and certainly needs work to make it more realistic;
however, we believe that it's a reasonable simple model to test our
solution to the second half of the reconstruction problem.

TKFST is not the main focus of the present paper -- it's just the
particular pairwise model we're using as an example.  We could have
used another pairwise model to illustrate transducer composition, and
we have attempted to make sure that the composition algorithm itself
is not presented in a TKFST-specific way.

> This time-dependent evolutionary model has essentially nothing to do
> with (is separable from) "transducer theory" as far as I understand;

We agree; with the exception that, if you take the time-dependent model as
representing a single branch, and consider extending it to an
arbitrary phylogenetic tree using the sum-product algorithm, then this
line of reasoning will lead you to an algorithm which is essentially
the same as the Inside-Outside algorithm on a Multi-SCFG constructed
by transducer composition.

For example, the multidimensional DP algorithm of Hein (PSB, 2001) for
computing the TKF model likelihood on a binary phylogenetic tree is
exactly equivalent to the a Forward algorithm for a composite
phylo-transducer. Though not originally expressed in transducer
formalism, Hein's algorithm is (arguably) easier to explain that way.

So there are conceptual relationships. It's also true that serial
composition of transducers amounts to "adding" two branch lengths, so
this is another relationship to the time-dependent theory (indeed, it
can be considered a "test" of the additive property of time-dependent
models).

But these are subtle connections. We agree with the reviewer that these
two questions (branch length and tree topology) can for most intents &
purposes be considered to be separable -- just as, with substitution
models, matrix exponentiation is separable from pruning.

> "transducer theory" is just a different set of words for conditional
> probability models, with the evolutionary continuous-time Markov
> model providing the particular _parameters_ that get plugged into
> the conditional probability model.
> (That is, I believe that I can interchange a
> "string transducer" with an HMM and "parse tree transducer" with an
> SCFG with no loss of information. If this is true, the exact names of
> things aren't important so long as they're clearly defined and
> explained for the particular purpose at hand, and one shouldn't be
> trying so hard to brand "transducer theory" as new. If I'm wrong, I
> would like to see the authors' clear explanation.)

While they are very similar, the essential difference is that there is
no general-purpose recipe to convert a Pair-HMM or Pair-SCFG into a
multiple-sequence HMM or SCFG. In order to do this you would need

(a) to label one sequence as the ancestor and the other as the
descendant;
(b) to ensure that the HMM/SCFG was then conditionally normalized;
(c) to devise an appropriate algorithm for constructing the state
space and the rule space of the composite machine (which essentially
amounts to co-ordinating transitions between the component pairwise
machines).

Steps (a) and (b) are trivial, but step (c) is nontrivial and is the
point of this paper. (The analogous algorithm for string transducers
is described in (Holmes, 2003).)

The reviewer is correct that, e.g., Pair HMMs and "string transducers"
are very similar, so that a conditionally-normalized Pair HMM is
essentially the same as a string transducer. This just amounts to a
restatement of steps (a) and (b); the innovation is in step (c).

We did not aim arbitrarily to brand transducers as new; rather, we are
attempting to highlight connections to existing literature.  We have
attempted to make this clearer in the revised manuscript by
emphasizing the term's origin in the computational linguistics
literature ("parse-tree transducer" comes from the same literature).
We have also attempted to make clearer the tight connections between
existing HMM theory in bioinformatics and the "transducer theory"
described in computational linguistics (c.f. Mehryar Mohri et al.)

The idea of transducer composition, above and beyond the
straightforward idea of a conditional probability model, has been used
in several other recent papers. For example:

(1) Paten B, Herrero J, Fitzgerald S, Beal K, Flicek P, Holmes I,
Birney E.  Genome-wide nucleotide-level mammalian ancestor
reconstruction.  Genome Res. 2008.

(2) Satija R, Pachter L, Hein J.  Combining statistical alignment and
phylogenetic footprinting to detect regulatory elements.
Bioinformatics. 2008 May 15;24(10):1236-42.

(3) Varadarajan A, Bradley RK, Holmes IH.  Tools for simulating
evolution of aligned genomic regions with integrated parameter
estimation.  Genome Biol. 2008 9(10):R147.

We think that the transducer formalism makes these papers considerably
simpler to read (and provides a connecting thread between them).

> As an example of
> what I mean by underemphasizing the evolutionary model, consider the
> very term "statistical alignment" that the authors use to describe for
> their subfield: the heart of the approach is neither statistical (a
> "statistic" being an arbitrary measure of something, as opposed to a
> probability model) nor really about "alignment" (the problem at hand
> is phylogenetic inference, not alignment; alignment is necessary, but
> only as a nuisance variable in the inferences that the authors are
> actually interested in making) -- not to mention that there is a large
> body of other work on probabilistic alignment approaches that the
> authors seem to be trying to marginalize by distinguishing only their
> particular favorite flavor as "statistical alignment".

We intended to use the term "statistical alignment" in the context
introduced by Jotun Hein's 2000 investigation of the TKF91 model.  It
has since been used to describe probabilistic alignment
approaches that attempt to optimize, and/or sample from, a
phylogenetic likelihood function based on an evolutionary model of
indels.

This phylogenetic likelihood function is our understanding of the
difference between "statistical alignment" and other probabilistic
approaches to multiple alignment -- such as aligning everything to a
consensus HMM (c.f. 'hmmalign' in the HMMER package), or combining
Pair-HMMs with a non-phylogenetic objective function for
posterior-decoding (c.f. PROBCONS and AMAP).

However, we agree with the reviewer that the term is ambiguous, and
after some consideration we have de-emphasized the term in the revised
manuscript, leaving only the following sentence:

  The literature on phylogenetic alignment using probabilistic indel
  models has been dubbed ``Statistical Alignment'' by Hein et al
  (2000).  However, we have avoided this term in the present
  discussion, since almost every structural RNA alignment program
  makes some use of statistics.

Other than our (now reduced) use of this phrase, we hope there is
nothing about this paper that implies we are trying to marginalize
other alignment approaches. In fact if there is, we would greatly
appreciate any specifics the reviewer can provide, as this is
something we'd seek to correct.

> I struggled, section by section, to decipher in the Results what the
> take home message of the paper -- or even any individual section -- is
> supposed to be. Each section of Results has similar problems; I'll do
> a fairly detailed analysis of just the first results section, then
> I'll be brief with the rest.
> 
> Results 2.1 attempts to describe an experiment in which the authors
> ask whether it is necessary to use "covariant" rate models, as opposed
> to point substitution rate models, to accurately reconstruct ancestral
> RNA sequences. There are immediately a number of problems just in the
> conception of this experiment, including:
> 
> 1. How is it relevant? The introduction claimed that "the central
> result of this paper is the development ["of a theory for statistical
> alignment and reconstruction of RNA"] building on the transducer
> framework for protein and DNA reconstruction". But rate matrices for
> single residues or even for base pairs are standard models in the
> field [...] and have nothing really to do with "transducer theory"
> nor with inference of indel structure, supposedly the main point of
> the paper.

We agree that this section distracts from the theoretical presentation
of transducer composition, and we have moved it to the supplementary
info.

To answer the question: "How is it relevant [to the central result]?"
We hope that this is primarily answered by the revisions we have made
to the paper; some further discussion follows here.

A phylo-SCFG such as the one we are benchmarking here can be viewed as
a special case of a multi-sequence SCFG, and can be similarly derived
using transducer composition. (If the parse-tree transducer does not
allow indels, i.e. if the indel rate is zero, and if its substitutions
use a time-dependent parameterization based on a continuous-time
Markov chain, then you end up with a phylo-SCFG like EVOFOLD or
PFOLD.)

So, in addition to being a framework for constructing multi-SCFGs with
indels and covariant substitutions, this is (at the same time) a
framework for constructing phylo-SCFGs. In case this seems like
excessive machinery for such a simple task, we note that there is
precedent for viewing phylo-HMMs as constrained or trivial special
cases of multi-sequence transducer-based HMMs (c.f. Siepel et al;
Diallo, Blanchette et al) and that such a view can be a useful guide
as to what approximations to make. So we think it is useful to make
this connection.

Multi-sequence SCFGs are memory-hungry. Pending development and
implementation of more efficient constraints or sampling algorithms
(such as the Redelings-Suchard MCMC kernel cited in the paper, which
involves doing two composite-transducer DP steps followed by a
Metropolis-Hastings accept/reject step), we are not yet able to do
systematic high-throughput tests of the more general & complex
multi-SCFGs (such as TKFST-based multi-SCFGs) which allow indels as
well as substitutions.

To recap, so far we only have a general-purpose transducer compositor
(that's this paper); this is a step towards building a
Redelings-Suchard sampler for RNA, but only a step; building the R-S
sampler (or some other efficient MCMC sampler) would be a significant
further step.

Efficient general-purpose implementations of phylo-SCFGs do currently
exist (conditional on an alignment) and xrate is one such general
implementation.

Our goal, in including the xrate reconstruction benchmark, was to
provide signposts to later multi-SCFG tests, by beginning with a
benchmark of the special-case phylo-SCFG.

However, we thank the reviewer for pointing out that these results
detract from the central message, and so we have accordingly moved
them to the supplementary material.

> (incidentally, the authors cited zero prior work on using base
> pair rate matrices was cited, but a fair amount exists)

The original paper did cite two papers by Knudsen and Hein
that use basepair rate matrices and the revised paper adds a citation
to RSEARCH (see above).  We can certainly expand on this with
further citations if the reviewer feels it appropriate.
We intended no slight to previous work, and have attempted to
emphasize prior art in the revised manuscript.

> 2. Why do the experiment at all, when the answer is already known and
> trivial? Obviously a point substitution model will not constrain
> ancestral base pairs to be valid base pairs; obviously modeling with
> base pairing correlations makes for a better model of a structural
> RNA. The reason to do a *quantitative* experiment to address a
> question whose qualitative answer is trivially obvious would be if the
> quantitative error rate was meaningful in some sense that had been set
> up in a problem definition statement. This would require deeper
> thought and discussion about how accurate an ancestral RNA
> reconstruction actually needs to be, for an intended purpose. For
> example, one might imagine a cost-benefit analysis of how many
> sequences need to be synthesized to find a functional ancestral
> sequence for a given error rate, if we're taking this research program
> seriously; then there might be a threshold prediction error rate that
> one needs to drop below, in order to make synthesis and testing of the
> predictions reasonable.

These are intriguing suggestions. The proposed Operations Research
study of the entire problem (including synthesis) is an especially
interesting new direction.

Our less-ambitious approach to the research program was to set up a
competitive benchmark of different reconstruction methods, using some
metric of reconstruction quality.

We believe that "percentage nucleotides correctly reconstructed" is a
reasonable benchmark metric for reconstruction quality.  It has been
used previously in other work. (See, for example, Paten et al,
Genome-wide nucleotide-level mammalian ancestor reconstruction. Genome
Res. 2008.)

We acknowledge that our current results in this area are preliminary,
which is why we have now relegated this section to the supplementary
info.

We are certainly trying to take this research program seriously.
However, these methods are challenging -- it took us three years to
get this far (i.e., implementing RNA structure on a three-branch tree)
-- and even our tightly-constrained algorithms push the limits of current
computational power, except on very simple RNAs.

Nevertheless, we think the transducer composition and loopy DP
algorithms, combined with our working implementations, are noteworthy
enough to publish. (For example, we think this is a useful step
towards a Redelings-Suchard kernel for RNA.)

> But instead, the authors do go ahead and draw
> only qualitative conclusions from their results in Fig 1, that
> modeling base pairs is better than not modeling base pairs: "we
> conclude that deep phylogenetic reconstructions of ribosomes, and
> other RNAs, will require covariant substitution models that take
> account of RNA secondary structure". But this, anyone working on RNA
> already "knew". Wait, the authors might say, maybe we shouldn't trust
> our intuition? Fine -- but ironically, in the very next sentence, they
> show how willing they are to make essentially the same obvious
> qualitative leap, on the most relevant point to the theory they want
> to express, without any quantitative analysis: "We may reasonably
> deduce that indel reconstructions will similarly need to take account
> of RNA structure". This "deduction" does not follow from the residue
> substitution results in Figure 1 any more than it already followed
> from what we already knew about RNA structure and evolution.

We agree, these particular qualitative conclusions were not 
surprising, and the logical chain of reasoning (placing these
benchmarks before the development of the model, when in fact they are
limited benchmarks OF the model) was muddled. We have removed these
parts.

We have every intention, eventually, to investigate the indel
reconstruction accuracy in a systematic high-throughput way (as we
have begun to investigate the accuracy of substitution
reconstruction). The memory complexity of the 3-sequence algorithm
means that most structures in RFAM are just out of range on current
hardware, so we restricted ourselves to tests of a simplified model.

Our revised manuscript discusses these limitations more clearly.

> Even beyond the conception and rationale of the experiment shown in
> section 2.1, the design and execution of the experiment are
> inadequately explained. What base pair rate models did you estimate --
> a completely unconstrained 16x16 base pair model? a general-reversible
> model? or a model with more constraints? What "75-taxon phylogeny" did
> you use (Rfam does not provide phylogenies); with what branch lengths?
> When you simulate 5000 "random alignments" of descendant sequences on
> this tree, how do you choose the initial ancestral sequence for each
> simulated alignment (it must have a base-paired structure for the
> experiment to be reasonable; there's many ways you could have done
> this, but it's unclear which, if any, were used).

In our revised manuscript, we have attempted to explain these points
more clearly for ease of reproducibility.

Most of the explanations that the reviewer is seeking were available
in the original manuscript, but as the reviewer points out, they were
frequently obscured.  We have attempted to correct this.

Specifically:
-- The actual distribution of ancestral structures is
unimportant (since we're not attempting to infer structure in this
particular benchmark, but are assuming that we already know the
correct structure) but we could have stated this more clearly.
-- Although we did say that the model was "reversible", we didn't
explicitly state that the 16x16 rate matrix was otherwise fully
general and unconstrained. Nor did we explicitly say that the
ancestral sequence was sampled from the equilibrium of this reversible
model.
-- The parametric structure of the 16x16 rate matrix and the initial
distribution of ancestral structures were both based on the PFOLD
phylo-grammar, which was referenced in the following benchmark
section, but not in this section.
-- The actual grammar over structures should not affect the final
result for this section, because we are assuming that we *know* the
true structure. The only thing that really matters is that the
structure includes both basepairs and single nucleotides, which we
then handle separately. Again, it would have been better to err on the
side of excessive experimental detail, rather than leave this out.

We hope that all of these explanations are now clearer in the text,
and we thank the reviewer for the feedback.

> Why did you strip only the ancestral (root) sequence out of the
> alignment? shouldn't _all_ internal node sequences be stripped out,
> if we're trying to simulate the ability to reconstruct ancestral
> seqs from extant seqs?

Yes, this _is_ what we did -- we took out all the internal node
sequences and the ancestral root sequence. By "ancestral sequences"
we didn't mean only the root sequence, but all the ancestors
(i.e. internal nodes) in the tree. We have attempted to clear this up
in the text.

> Results 2.2 follows a similar pattern. The authors use bold face to
> "emphasize that this [TKFST] is not just a method for simultaneous
> alignment and folding of RNA" -- even before they make a concrete
> problem definition of what TKFST _is_! A reproducible description of
> TKFST isn't given here; it is "reviewed in Section 3.1". No, it isn't
> given there either; Section 3.1 meanders in an unclear fashion, still
> without problem definition, hitting a variety of topics connected to
> TKFST, while saying at various points things like "full details... are
> given in the Supplementary Material". So much for this "theory" for
> "computational reconstruction of the RNA World" being the "central
> result" of the paper! What the heck is it doing lying around in pieces
> in the Results, Methods, and Supplementary Material, as opposed to one
> clear central exposition?

The reason for this apparent omission of a full description of TKFST
is that is not the central result of this paper.
Rather, the "theory" we were referring to was the
composition algorithm for extending any pairwise SCFG to multiple
sequences related by a tree. TKFST was merely an example pairwise
model that we were using to illustrate this algorithm.

(At the risk of excessively repeating ourselves: TKFST is a prototype
solution to the "time-dependent parameterization" problem, a.k.a. the
"branch length" problem; whereas the problem we're really trying to
solve here is the "multiple sequences" problem, a.k.a. the "tree
topology" problem.)

Hopefully the increased emphasis on the transducer composition
algorithm helps clear this up.  We have tried to give a slightly more
concrete description of the TKFST model; we can beef this up if the
reviewer feels it appropriate.

In order to emphasize that the transducer composition algorithm is our
main result, we have moved the results which the reviewer refers to to
the supplementary material.  These results did not use transducer
composition; they just used TKFST as a Pair-SCFG to do progressive
alignment (or, strictly speaking, multiple alignment by single-linkage
clustering).

We acknowledge that including this benchmark in a paper about
phylogenetic transducer composition was, therefore, rather confusing.
The reason we included it at all was, again, that this alignment
quality benchmark does resemble one of the benchmarks we intend to
repeat with transducer-based algorithms as soon as (a) RAM becomes
cheap enough or (b) we develop more efficient/constrained algorithms
by extending the techniques presented here.

Furthermore, this benchmark does speak to the usefulness of the TKFST
model, which (while not the central result) *is* relevant to this
paper, since we have used TKFST as our "test model" for transducer
composition.

As with the other benchmark, we have moved this section to the
supplementary info, and we hope the rewrite makes things clearer.

> My comments on the rest of the paper would continue repetitively in
> the same vein but since this review is already long, I'll stop. In
> summary, I think the authors are working on very interesting issues
> that they're explaining atrociously. Writing a clear paper requires
> more discipline. The authors need to go back to the drawing board for
> this paper, pick one point they want to make, and write a clear,
> logical paper that conveys one nice story rigorously.

Thanks again for the constructive criticism. We have tried to follow
these suggestions.

> some small points:
> 
> - group II introns are not "76-91 nt" (Table 3 and elsewhere); they
>   are much larger than this. The authors must be using the Rfam
>   Intron_gpII alignment, which only deals with two of the six
>   consensus structural domains (domains V and VI).

Thank you for this correction. We have added text noting this at
several places in the manuscript and supplementary information. (The
sequences were taken from BraliBaseII and thence from Rfam 5.0.)

> - it's a lot more useful to a reviewer to get a manuscript in
>   manuscript form (i.e. at least 1.5x linespaced if not doublespaced),
>   not a tightly spaced journal format, so there's ample room for
>   annotation.
>
> - the fonts in Figures 1 and 2 are too small.

We have implemented these suggestions.

> Reviewer #2 (Remarks for the Author):
> 
> I. The manuscript describes an method for simultaneously inferring the
> alignment and secondary structure of three modern RNA sequences along
> with that of the internal node on a 3-taxon tree - a previously
> unsolved problem to the extent of my knowledge. The manuscript also
> assesses the TKF Structure Tree (TKFST) model of RNA structure
> evolution in terms of its ability to help reconstruct ancestral
> alignments and structures. Both of these results (new method, model
> assessment) are important steps towards the goal of reconstructing the
> sequence and structure of ancestral RNA molecules, which is a very
> interesting biological question.
> 
> This topic is by nature complex and involved, and I am unfortunately
> not an expert in RNA structure. Nevertheless, I think that the authors
> display a deep and intuitive grasp of this field that is quite
> impressive, while being moderately understandable to the outsider. In
> most cases, I was able to follow the description of rather complex
> procedures on my first pass through the paper even though I was not
> previously acquainted with SCFGs in any detail. I think that the
> readability for readers that are less familiar with the background can
> be improved by adding some background in a few places, which I will
> try to indicate below.
> 
> II. My major advice on improving the readability of the paper is to
> decrease or remove emphasis on the names of programs. In general, an
> inference or estimate should be described in terms of (i) the model
> used, (ii) the type of estimate (e.g. maximum likelihood estimate,
> posterior mean, etc.) and (iii) the algorithms employed. It should not
> be described in terms of the name of the program that is run to
> perform the inference. For example, in the beginning of section 2.1,
> the authors write "We first used xrate to estimate parameters..." when
> they should say something like "We used an EM algorithm to compute
> maximum likelihood estimates of parameters for the substitution
> model. We held the alignment fixed, and summed over the unknown
> secondary structure and ancestral sequences."

This is an excellent point.  We have modified the text accordingly.

> (I recognize that the authors wish to point out their work in
> extending existing programs to provide infrastructure for estimating
> ancestral sequences.)
> 
> Secondly, I think that the authors downplay the very real value of
> their analyses by focusing too heavily on the exact protocol of the
> analyses and playing down the questions and results in addition to the
> methods. For example, in Section 2.1 they write "We implemented ..." ,
> "We first used xrate ..." , "Next we used the companion program
> SIMGRAM ...", "We stripped the ancestral sequence..." It would be
> helpful if the authors began section 2.1 with a description of the
> tests that would be performed in that section, and what questions they
> answer, instead of beginning by describing xrate. Similarly, in
> section 2.2 the authors make an important point in bold font: "We
> emphasize that this is not just a method for simultaneous alignment
> and folding of RNA. Our intention here is a critical assessment of the
> realism of TKFST as an evolutionary model, not simply an
> implementation of the algorithm of Sankoff."  Nevertheless, this
> section is titled "Multiple Alignment and RNA folding". Hopefully
> these examples clarify my general point about how the authors can
> improve readability.

Thank you for the suggestion.  We have moved what were Sections 2.1
and 2.2 to the supplementary material in order to make clear that our
principal result is the transducer composition algorithm for building
models of multiple RNAs on a tree.

> III. Specific suggestions
> 
> 1. [ Abstract ] I presume that xrate estimates maximum likelihood
> parameters estimates using the EM algorithm. I presume that xrate sums
> over all structures when estimating parameters? When it is used to
> estimate structures, I presume that structure estimates are
> conditioned on maximum likelihood rate estimates?
> 
> These points should be clarified somewhere, though perhaps not in the
> abstract.

This is correct: structure estimates are conditioned on the "model,"
which is parametrized with ML rate estimates.  We have moved the
"xrate" results to the supplementary material (so they aren't
mentioned in the abstract).  We have also clarified this point in the
supplementary material.

> 2. [Abstract] The abstract is written as if the main topic of interest
> is programs, not algorithms and biological questions. However, the
> authors underestimate the importance of their work by focusing on
> programs.

We have rewritten the abstract accordingly to place more focus on the
algorithms rather than the programs which implement them.

> 3. [ pg 2. / Section 2.1 ]
> 
> While implementing programs is an important part of biological
> research, I think that the authors are downplaying their true
> contribution in this section by describing the program names that were
> used instead of focusing on the methods and algorithms.

We agree and have revised the text accordingly.

> [pg 2 / Section 2.1 ] "We implemented ancestral sequence
> reconstruction in xrate, including posterior probabilities that the
> reconstructions are correct."
> 
> This single sentence should be expanded to at least one or two
> paragraphs describing the type of estimation that is being performed
> as well as the type of algorithm that implements this estimation. I
> presume that (i) xrate estimates parameter values of the substitution
> model using maximum likelihood.  (ii) I presume that in the definition
> of the likelihood the alignment is held constant, but the secondary
> structures and ancestral sequences are summed over (as latent
> variables).  (iii) The maximization over parameters is carried out
> using an EM approach.
> 
> If this is the case, then what kind of estimates of the structure, and
> of ancestral sequences, are being performed? I presume that xrate
> conditions on the maximum likelihood estimates of substitution
> parameter values and then selects the ancestral sequence (or
> structure) that has the highest posterior probability? If so, then
> these points might be a good starting place for a description of xrate
> and the extension that you performed.

We have expanded this text in the Supplementary Material.

> 4. [ pg 3 / Section 2.2 ] The authors should retitle this section to
> indicate that it is primarily about the critical assessment of the
> TKFST model, and focus less on the exact protocol. Perhaps the exact
> steps could be described in a separate "Methods" section that
> describes protocol details. These details must be provided so that the
> reader fully understands what the authors are doing, but they are not
> usually the main point.

This is an excellent suggestion.  We have attempted to make clear that
these results form a critical assessment of TKFST's appropriateness as
a model of RNA structural evolution when we reference these results in
the main text.

> The authors explicitly say "We emphasize that this is not just a
> method for simultaneous alignment and folding of RNA. Our intention
> here is a critical assessment of the realism of the TKFST as an
> evolutionary model." This is a good point - but the title and logical
> structure of this section focuses more on the protocol details for
> running programs.
> 
> 5. [pg 3 / "pair SCFG" ] This has not been defined yet. Many readers
> will be unfamiliar enough with SCFGs that it is important to define
> pair-SCFG's before saying too much about them.

The first use of "Pair SCFG" now appears after SCFGs have been
defined, and in the context of a discussion of pairwise models, while
citing previous work. Hopefully these pointers are now enough.  We can
add more information if the reviewer feels that it would be helpful.

> 6. [ pg. 3 ] It would be helpful to add one or two more sentences of
> explanation about sensitivity and PPV mean, and how they are applied
> to alignments and RNA structures.

We have added explanatory text to the table caption, defining and
discussing sensitivity and PPV for alignments and structures.

> 7. [ pg. 4 ] It is curious that the stemloc native grammar doesn't
> seem to cause much improvement in alignment sensitivity and PPV even
> though it has an affine gap penalty. Can you say why this is the case?
> What is the average gap length - is it near 1?
>
> 8. It would be useful to give parameter estimates from PFOLD or the
> stemloc-native grammar to indicate what kind of properties these RNA
> molecules have, and whether or not a model like TKFST that has fewer
> parameters should be expected to do a good job modeling the evolution
> of such sequences. For example, if the average gap length is near
> zero, then that parameter is not helping the PFOLD model much, but if
> it is large, then we should expect the fewer parameters of the TKFST
> model to result in worse model fit and inferior prediction quality.

The reviewer's insight into the affine gap penalty appears to be very
accurate. We have added the following sentences to the discussion of
these results:

One possible (partial) explanation [for the close results] is that the
affine gap penalty is not helping all that much: the mean gap length
in the stemloc grammar ranges from 1.3 to 3.5 bases (in loops) and 1.4
to 1.7 basepairs (in stems), so the improvement over a linear gap
penalty is small. (The mean gap lengths are specified as ranges
because stemloc's native grammar set includes four distance settings;
mean gap length is a function of distance.)

> 9. [pg. 5 ] "To test this, we tried re-folding the multiple alignments
> with xrate..."  This sentence is unclear because it uses the program
> name xrate instead of describing the procedure that was carried out,
> and because the word "re-folding" is a bit too terse. You might say,
> "To test this, we re-estimated rate parameters and secondary structure
> using the PFOLD grammar while fixing the multiple sequence alignment
> to the value estimated in the previous step." You might then
> optionally add that this was accomplished using xrate if you wish to
> emphasize which program was used.

We have replaced this sentence with the one suggested by the reviewer.

> 10a. [pg. 5] "The structure prediction accuracy of the TKFST and
> stemloc-native grammars was much better when combined with xrate, and
> both grammars were comparable..."
> 
> The authors should not say "when combined with xrate" but should
> indicate what they are using xrate to do.

We have changed this to "...when combined with a post-processing
structure prediction step".

> 10b. When the authors re-estimate the structure using xrate, they use
> a different model (PFOLD) to do so. Why, then, are these new structure
> estimates taken to represent "the structure prediction accuracy of the
> TKFST and stemloc-native grammars"? Don't they represent the
> structure-prediction accuracy of the PFOLD grammar, conditioned on
> alignments inferred using the TKFST and stemloc-native grammars?

Yes, this is true, so we have deleted the words "of the TKFST and
stemloc-native grammars", and added the reviewer's suggested phrasing
to the table caption.

> I don't see a good reason why the PFOLD model should be used during
> re-estimation of the secondary structure. If the problem is early
> commitment to incorrect second structures during an algorithm like
> progressive alignment, then re-estimating the structure with the TKFST
> and stemloc-native models (respectively) should eliminate the problem.

This is true, but we currently have no implemented software that can
predict structures from multiple alignments using parse-tree
transducer models or multi-SCFGs. We only have software (xrate) that
can do this for phylo-SCFGs (like PFOLD). Ideally we would use the
same model throughout the analysis; however the point here is to
demonstrate that the alignments are at least sufficiently accurate to
get reasonable structure predictions from a method that takes an
alignment as input (PFOLD).

> 10c. [Table 2] I don't think the last two lines should be
> characterized by "+xrate" since xrate is just a software
> package. Instead, you should choose a label that indicates what you
> are doing.  (I) I am not sure, but it looks like you are using PFOLD?
> (II) If so, then perhaps the last two rows should be written as
> "+PFOLD?"

True, we have now changed this from "+xrate" to "+PFOLD".

> 10d. [pg. 5] "We parametrized the singlet and branch transducers of
> the TKFST model using estimates reported by PFOLD, a phylo-grammar for
> RNA secondary structure prediction, and evoldoer, an implementation of
> the Structure Tree model for two sequences."
> 
> (i) How can you obtain estimates for the parameters in the TKFST model
> by using a different model - the PFOLD model?  (ii) This sentance
> again describes the names of program used instead of the methods and
> algorithms by which the estimates were constructed.  (iii) I would
> suggest rephrasing the last phrase "an implementation of the Structure
> Tree model for two sequences". This is because technically you don't
> implement models, you implement inference algorithms that use
> models. So, here, evoldoer implements inference of the pairwise
> alignment between two sequences under the TKFST model (I think). If
> one algorithm (DP) implements the hard part of a wide variety of
> inference methods then you could say so, I suppose.

(i,ii)
Thank you for pointing this out, and we apologize for the oversight.
We have described the origin of the so-called "PFOLD" parameters,
which we used because they have proven effective for RNA structure
prediction, in much more detail.  We used the substitution rate
estimates reported by Knudsen & Hein directly for our model.

(iii)
We have changed the text as recommended.

> 13. "Pair-SCFG" Can the authors describe a pair-SCFG in more detail?
> Although the manuscript compares the pair-SCFG to a pair-HMM, the
> authors should still point out (for example) that a pair-SCFG is an
> SCFG that emits pairs of letters, not a pair of SCFGs. It might also
> be worth explicitly pointing out that the transformation rule "S1 → x
> S2" in the SCFG corresponds to (in a pair-HMM) (I)state S1 emits
> letter "x" (II)state S1 then transitions to state S2.  These two
> things are probably worth pointing out, given that they do not add
> much length to the paper.

We have included this suggestion.  The introduction to the Methods
section describes the phenomena which Pair SCFGs can model and the
caption to Table 2 includes the example recommended by the reviewer.

> 14. [pg. 8] Define "X-terminals" and "Y-terminals." This is important
> terminology that is re-used, and it would be worth making sure the
> reader understands it here in order to clarify what follows. Also, you
> might clarify that X here represents the ancestor and Y the descendent
> along a branch of the tree, since this isn't always the case (I
> think).

We have replaced this terminology with just "terminals" for clarity and
we describe terminal symbols as analogous to emitted characters.
The text now tries to make clear that X is the ancestor and Y the descendant.

> 15. [pg. 10 ] "In practice, we find the size of the state space is
> approximately linear in the number of branches..."  Can you say
> whether you find this to be true for several different grammars, or
> just for (e.g.) the TKFST grammar? It would seem that the number of
> states should grow exponentially in the number of wait states for each
> branch transducer, or faster. However, because the TKFST model does
> not have affine gap penalties, it may have only one wait state,
> leading to the linear growth that you observed. Regardless of whether
> this guess is correct or incorrect, a comment about how affine gap
> penalties affect the number of states would be helpful here.

These are excellent points.

We do not have a proof of this complexity behavior, but we have observed this
behavior for several models.  However, as we do not have theoretical
upper-bound results, we have toned down these statements in the
text to clarify that they are merely empirical observations and that
we are unsure how they will generalize.

For technical reasons, TKFST actually has several wait states.
We believe that the growth will be sub-exponential in the number of
wait states because we are able to eliminate many of them using the
procedure described for eliminating "windback" states in the section
"Reduction of computational complexity."  We have noted this in the text.

> If this is the case, then could you say something like "In the case of
> the TKFST model" instead of "In practice..."? Also, can you clarify
> "realistic biological models never reach this bound"? What aspect of
> realistic models prevents the bound from being reached?

This bound required un-biological state configurations such as ones
which align un-related sequence.  We now state this clearly in the text.

> 16. [pg. 11] "... is sufficient for a likelihood-sampling algorithm
> over a tree".  The phrase "likelihood-sampling" is vague - technically
> the likelihood itself isn't being sampled. You could say something
> like "is sufficient for a calculating the likelihood of a full tree,
> which would be used in likelihood-based algorithms for sampling branch
> lengths, substitution rates, and other model parameters". Or perhaps
> you mean, sufficient for sampling the alignment and the RNA structure
> at all nodes?

We have changed this as recommended.

> [pg. 11] "Evaluation of the likelihood function of a
> node conditioned on its neighborhood..."  Can you rewrite this for
> accuracy and clarity?  (I) "... the likelihood function of a node ..."
> - this is vague, and focuses more on commonly used algorithms than on
> what exactly those algorithms calculate.

We have changed this as recommended.

> (II) "... conditioned on its
> neighborhood..." - technically, you are not statistically conditioning
> on the sequence at every node in the neighborhood, but just on
> (perhaps) the sequence at the ancestor node. The sequences at the
> other nodes are given, but you are acutally computing the probability
> of the sequences at these nodes, not conditioning on them.

We may not fully understand the distinction which the reviewer is
making, but if the sequence at a neighboring node of the node in
question is observed, then considering the distribution which is joint
with the observed neighboring node and the distribution which is
conditional w.r.t. the same node is equivalent.  The node is observed,
so they differ only by a multiplicative constant, ie, the marginal
evaluated at the observation.

We can state this point more clearly in the text if the reviewer feels
it appropriate.

> [pg 6., Fig 3] "Maximum likelihood reconstruction of ancestral
> structures" When you say "ancestral structures", do you mean the
> structure at the internal node on an unrooted 3-taxon tree? This is
> not the ancestor of the three leaf sequences, but the common ancestor
> of two of the sequences, using the third as an
> outgroup. Reconstructing the ancestor of all three would require
> knowing where the root of the tree was, if I understand correctly.

Thank you; this is an excellent point.  We have clarified the text
accordingly (in the caption of Figure 2 showing the corresponding
trees and inferred alignments).

> In summary, I hope that these suggestions are useful to the authors in
> improving readability of the paper. The only point that may be a real
> error is point 10a,b,c,d. Overall the manuscript was well-written and
> makes important contributions to the field of RNA structure
> reconstruction.
> 
> -Benjamin Redelings
>