### 2009-04-18 cover-letter-resubmission-1.tex

\documentclass[10pt]{letter}
\pagestyle{plain}
%\usepackage[ansinew]{inputenc}

%\addtolength{\topmargin}{-0.5in}

%\sloppy
%\nonfrenchspacing

\renewcommand{\baselinestretch}{1.0}

\address{
Ian Holmes\\
Department of Bioengineering\\
University of California\\
Berkeley, CA 94720\\
}

\signature{}
\begin{document}
\begin{letter}{}

\opening{Dear Prof. Stormo:}

\vspace{\baselineskip}

Thank you very much for your attention to our earlier manuscript and for forwarding the extremely encouraging reviews.

Based on the insightful comments of the reviewers, we have made sweeping changes to the structure and presentation of the manuscript.
We believe that it is greatly improved as a result.

We would be most grateful if you would consider the accompanying revised manuscript, along with the attached point-by-point response to the reviewers' earlier criticisms,
describing the changes we have made.

All authors have read and approved of this manuscript.

%As editor, we suggest David Haussler.

%We suggest the following reviewers:
%David Haussler (if not editor),
%Jeff Thorne,
%Marc Suchard,
%Jotun Hein,
%Sean Eddy.
%We have provided contact details for these reviewers in the online
%submission process.

\closing{Sincerely,\vskip 20pt Robert K. Bradley\\Ian Holmes}

%% If you have things to enclose, specify here
%\encl{Encl. 1}
%\encl{Encl. 2}

\end{letter}

\end{document}