### 2009-06-04 s3.tex
renamed software appendix (Text S3) to Text S4; added new Text S3 for TKFST grammars; added a bunch of other stuff

\documentclass[10pt]{article}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

% Text layout
\topmargin 0.0cm
\oddsidemargin 0.5cm
\evensidemargin 0.5cm
\textwidth 16cm 
\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Use the PLoS provided bibtex style
\bibliographystyle{PLoS2009}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **
%% Please insert a running head of 30 characters or less.  
%% Include it twice, once between each set of braces
\markboth{Evolutionary modeling}{Evolutionary modeling}

%% ** EDIT HERE **
%% PLEASE INCLUDE ALL MACROS BELOW
\input{defs.tex}
% no idea why \outside was redefined with a "\right," instead of a normal comma in defs.tex, but am redefining it back here - IH, 5/29/09
\renewcommand{\outside}[7]{\beta_{#1}\left(#2,#3,#4\right)}
%% END MACROS SECTION

\begin{document}

% Title must be 150 words or less
\begin{flushleft}
  {\Large
    \textbf{\titlestring: \supptext{3}}
  }
\\
\authorstring
\end{flushleft}

\newpage
\section{Inference grammars for the TKFST model}
\seclabel{TKFST}

This section describes multi-sequence grammars that can be used to infer alignments, structures and ancestors under the
TKF Structure Tree (TKFST) model.

\input{tkfst.tex}

\newpage
\section{The TKF Structure Tree model as a transducer} \seclabel{tkfst}

We can cast the TKF Structure Tree (TKFST) model as a transducer.
This means we can automatically deduce rules like those shown for the TKFST model in \secref{TKFST}.

\subsection{The single-sequence TKFST model as a singlet transducer}

The states of the singlet transducer are shown in \tabref{tkfstsinglet}.
Allowed transitions between these states are shown in \secref{TKFST}.

\begin{table}[!ht]
  \centering
  \begin{tabular}{lrrrr}
    state & $\type$ & $\absorb$ & $\emit$ & $e(\,\bullet\,|\mathrm{TKFST})$ \\ \hline
    $L$ & $\Sstart$ \\
    $I_L$ & $\Sinsert$ & & $(x,\Tnull)$ & $p(x)$ \\
    \\
    $S$ & $\Sstart$ \\
    $I_S$ & $\Sinsert$ & & $(x,y)$ & $p(x,y)$ \\
    \\
    $B$ & $\Sinsert$ & & $(L\,S)$ & 1 \\
  \end{tabular}
  \caption{
    \tablabel{tkfstsinglet}
    States of the singlet transducer of the TKF Structure Tree model \cite{Holmes2004}.
    Singlet transducers can only have states of type $\Sstart$ or $\Sinsert$.
    This is the indiegram-style transducer equivalent of the singlet TKFST SCFG in \secref{TKFST}.
  }
\end{table}

\subsection{The two-sequence TKFST model as a branch transducer}

The states of the branch transducer are shown in \tabref{tkfstbranch}.
Allowed transitions between these states are shown in \secref{TKFST}.

\begin{table}[!ht]
  \centering
  \begin{tabular}{lrrrr}
    state & $\type$ & $\absorb$ & $\emit$ & $e(\,\bullet\,|\mathrm{TKFST})$ \\ \hline
    $L$ & $\Sstart$ \\
    $I_L$ & $\Sinsert$ & & $(u,\Tnull)$ & $p(u)$ \\
    $M_L$ & $\Smatch$ & $(x,\Tnull)$ & $(u,\Tnull)$ & $p(u|x)$ \\
    $D_L$ & $\Smatch$ & $(x,\Tnull)$ & $(\Tgap,\Tnull)$ & 1 \\
    $W_L$ & $\Swait$ \\
    \\
    $S$ & $\Sstart$ \\
    $I_S$ & $\Sinsert$ & & $(u,v)$ & $p(u,v)$ \\
    $M_S$ & $\Smatch$ & $(x,y)$ & $(u,v)$ & $p(u,v|x,y)$ \\
    $D_S$ & $\Smatch$ & $(x,y)$ & $(\Tgap,\Tgap)$ & 1 \\
    $W_S$ & $\Swait$ \\
    \\
    $B_i$ & $\Sinsert$ & & $(L_i\,S_i)$ & 1 \\
    $B$ & $\Smatch$ & $(L\,S)$ & $(L\,S)$ & 1 \\
    $B_p$ & $\Smatch$ & $(L\,S)$ & $(L\,\Send)$ & 1 \\
    $B_e$ & $\Smatch$ & $(L\,\Send)$ & $(L\,\Send)$ & 1 \\
  \end{tabular}
  \caption{\tablabel{tkfstbranch}
    States of the branch transducer of the TKF Structure Tree model \cite{Holmes2004}.
    States which have the same names as states of the singlet transducer in \tabref{tkfstsinglet}
    are the branch equivalent of the corresponding singlet states
    (e.g. a $\Smatch$ state might be the branch equivalent of an $\Sinsert$ state).
    States $L_i$ and $S_i$ are the $\Sstart$ states of a sub-model (not shown) identical
    in structure to the singlet transducer.  They are used to insert a new stem-loop structure.
    This is the indiegram-style transducer equivalent of the conditional pair TKFST SCFG in \secref{TKFST}.
  }
\end{table}

\subsection{The multi-sequence TKFST model as a composite transducer}

We can use the state graph construction algorithm described 
in the paper and detailed in Text S1 to create a model of the simultaneous evolution
of several sequences.

\begin{figure}[!htb]
  \centering
  \includegraphics [scale=0.4] {figs/parent.pdf}
  \caption{\figlabel{parent}
    A simple example of transducer composition to build a multi-sequence model of two extant sequences.
    An ancestral sequence $W$ evolves into two descendant sequences $X$ and $Y$.
    A singlet transducer (the horizontal gray box) emit ancestral sequence and structure
    and two branch transducers (the gray boxes labeled $\Delta T_X$ and $\Delta T_Y$)
    mutate it according to the specified multi-sequence model.
    Gray nodes correspond to observed data and white nodes unobserved data.
  }
\end{figure}

Consider the simple model shown in \figref{parent}.
The state of the multi-sequence model describing this model is a 3-vector $\bvec{a} = \left( a_1,\,a_2,\,a_3 \right)$,
where $a_1$ is the state of the singlet transducer generating the ancestral sequence $W$
and $a_2$ and $a_3$ are the states of the branch transducers evolving $W$ into
extant sequences $X$ and $Y$.

We can show some of the allowed transitions of this multi-sequence model.
The state of the branch transducer associated with the active node $n$ is shown in bold.

\paragraph{Stem creation.}

A stem is created at the root node $W$ (corresponding to a bifurcation in the singlet transducer $a_1$)
and survives in the sequence $X$ at node 2 but is deleted in the sequence $Y$ at node 3.

\begin{align}
  \nakedthreevec{1:}{2:}{3:} \quad
  \threevec{L}{L}{\mathbf{L}} &
  \to \threevec{\mathbf{L}}{W_L}{W_L}
  \to \threevec{B}{B}{B_p}
  \to \threevec{L}{L}{\mathbf{L}} \, \threevec{S}{\mathbf{S}}{\Send}
  \to \threevec{L}{L}{\mathbf{L}} \, \threevec{\mathbf{S}}{W_S}{\Send} \\
  & \to \threevec{\mathbf{L}}{W_L}{W_L} \, \threevec{\mathbf{S}}{W_S}{\Send}
\end{align}

\paragraph{Stem insertion.}

A stem sequence is inserted in sequence $X$ at node 2.

\begin{align}
  \nakedthreevec{1:}{2:}{3:} \quad
  \threevec{L}{L}{\mathbf{L}} \to \threevec{L}{\mathbf{L}}{W_L}
  \to \threevec{L}{B_i}{B}
  \to \threevec{L}{L}{\mathbf{L}} \, \threevec{\Send}{S}{\mathbf{S}}
  \to \threevec{\mathbf{L}}{W_L}{W_L} \, \threevec{\Send}{\mathbf{S}}{W_S}
\end{align}

\paragraph{Stem termination.}

All stem sequences are ended by (possibly empty) loop sequences.

\begin{align}
  \nakedthreevec{1:}{2:}{3:} \quad
  \threevec{\mathbf{S}}{W_S}{W_S}
  \to \threevec{B_e}{B_e}{B_e}
  \to \threevec{\Send}{\Send}{\Send} \, \threevec{L}{L}{\mathbf{L}}
  \to \threevec{\Send}{\Send}{\Send} \, \threevec{\Send}{\Send}{\Send}
\end{align}

The functions $\alpha(t)$, $\beta(t)$ and $\gamma(t)$ are parametrized by the insertion and deletion rates of the Structure Tree model.
They are defined for loop sequences as
\begin{align}
  \kappa_1 &= \lambda_1 / \mu_1 \\
  \alpha_1 &= \exp \left( -\mu_1 t \right) \\
  \beta_1 &= \frac{\lambda_1 \left( 1 - \exp \left((\lambda_1 - \mu_1) t \right) \right)}{\mu_1 - \lambda_1 \exp \left( (\lambda_1 - \mu_1) t \right) } \\
  \gamma_1 &= 1 - \frac{\mu_1 \left( 1 - \exp \left((\lambda_1 - \mu_1) t \right) \right)}{\left( 1 - \exp (- \mu_1 t) \right) \left(\mu_1 - \lambda_1 \exp \left( (\lambda_1 - \mu_1) t \right) \right) }
\end{align}
and similarly for stem sequences \cite{Holmes2004}.